---
title: "bootstrap"
author: "Matt Klapman"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(here)
library(rsample)
devtools::load_all()
```

```{r}
true_result <- 53.1-45.2
rcp_midterms <- read_csv(here("DATA","RCP_midterms.csv"))
rcp_midterms <- rcp_midterms %>% mutate(prop_diff = (Dem - Rep)/100, ord = 1:length(prop_diff),
                                        Dates = as.Date(Dates, "%m/%d/%Y", 
                                                        origin = as.Date("01/01/2017","%m/%d/%Y")),
                                        #tr_num_dates = trunc(as.numeric(Dates) - 17167, prec = -1)/10
                                        num_dates = as.numeric(Dates) - 17167,
                                        prob = num_dates/sum(num_dates),
                                        rel_Dem = Dem/(Dem + Rep), 
                                        rel_Rep = Rep/(Dem + Rep), 
                                        rel_diff = rel_Dem - rel_Rep,
                                        prob_full = double(length(prob)))

rcp_midterms$prob_full[rcp_midterms$SampType == "LV"] <-
  rcp_midterms$prob[rcp_midterms$SampType == "LV"] + 1/3/309

rcp_midterms$prob_full[rcp_midterms$SampType == "RV"] <-
  rcp_midterms$prob[rcp_midterms$SampType == "RV"] + 2/9/309

rcp_midterms$prob_full[rcp_midterms$SampType == "A"] <-
  rcp_midterms$prob[rcp_midterms$SampType == "A"] + 1/9/309
```

```{r}
prop_diff_est <- mean(rcp_midterms$prop_diff)


# Bootstrap procedure with no weights
# ntimes <- 1000
# boot_propdiff <- rerun(ntimes, sample(rcp_midterms$prop_diff, 
#                                       replace = TRUE))
# boot_dist <- map(boot_propdiff, mean) %>% flatten_dbl
# 
# sd(boot_dist)
# 
# ci_noweight <- quantile(boot_dist, c(.025, .975))
```


I then started to do a basic bootstrap to simulate what different versions of an election would be, considering each poll to be a simulation of an election. But, I would like to place some type of wheighting or a probability vector to figure out how likely each poll could be selected. I think it would be best to weight by how recent the poll was taken before the election. I also noticed that I am leaving out valuable information by only bootstrapping the difference in proportions, as there is still the effect of the undecided voter. The difference between each party could be large, but if both only got about 30% of the vote, then there could be a decent number of undecideds.

So, I think I should use the `bootstraps()` command from the rsamples package to bootstrap entire rows from the dataset. 


```{r}
# Bootstrapping with weights

# boot_propdiff <- rerun(ntimes, sample(rcp_midterms$prop_diff, 
#                                       replace = TRUE,
#                                       prob = rcp_midterms$num_dates/sum(rcp_midterms$num_dates)))
# 
# boot_mean <- map(boot_propdiff, mean) %>% flatten_dbl
# mean(boot_mean)
# sd(boot_mean)
# ci_weight <- quantile(boot_mean, c(.025, .975))


w_est <- with(rcp_midterms, weighted.mean(prop_diff, w = prob)) # weighted estimate of proportion.


# Bootstrapping with keeping original percentages. 
new_boot <- bootstraps(rcp_midterms, times = 1000)

new_boot <- new_boot %>% 
  mutate(
    mean_diff = map_dbl(splits, 
                        ~ mean(analysis(.)$Dem - analysis(.)$Rep))
  )

ci_orig_perc <- quantile(new_boot$mean_diff, c(.025, .975))/100



```



After looking into some more literature, I found that a common strategy for those doing analysis on polls is to quantify the amount of support by dividing by the total of Democrats and Republicans. This gets rid of the undecided voters and third party voters. So, I will try to bootstrap with this new variable of difference in relative support. 

```{r}
relative_boot <- bootstraps(rcp_midterms, times = 1000)

relative_boot <- relative_boot %>% 
  mutate(
    mean_rel_diff = map_dbl(splits, 
                            ~ mean(analysis(.)$rel_diff))
  )
rel_est <- mean(rcp_midterms$rel_diff)

sd(relative_boot$mean_rel_diff)

ci_relative <- quantile(relative_boot$mean_rel_diff, c(.025, .975))
```

This turns out to be not too great of an estimate, even further bringing us from a realistic estimate. Though this is more similar to the models of other data scientists in the field.

What if I bootstrap just subsetting out polls of likely voters, which should hopefully give the most accurate results.

```{r}
ntimes = 1000
rcp_lv <- filter(rcp_midterms, SampType == "LV")

boot_lv <- rerun(ntimes, with(rcp_lv,sample(prop_diff, 
                                            replace = TRUE,
                                            prob = num_dates/sum(num_dates))))

boot_lv_dist <- map(boot_lv, mean) %>% flatten_dbl
lv_est <- with(rcp_lv, weighted.mean(prop_diff, num_dates/sum(num_dates)))
sd(boot_lv_dist)
ci_lv <- quantile(boot_lv_dist, c(.025, .975))



```


I have done some type of bootstrap a few times now, while only making slight changes. I should write a bootstrap function to call which will output a confidence interval.

```{r}
ci_lv <- with(rcp_lv, boot_ci_propdiff(prop_diff, prob = prob))
ci_weight <- with(rcp_midterms, boot_ci_propdiff(prop_diff, prob = prob))
ci_noweight <- with(rcp_midterms, boot_ci_propdiff(prop_diff))
ci_relative <- with(rcp_midterms, boot_ci_propdiff(rel_diff))


confints <- as.tibble(rbind(ci_noweight, ci_weight, ci_relative, ci_lv)) %>% 
  add_column(BootType = c("No Weights", "Weights", "Relative", "LikelyVoters")) %>%
  add_column(Est = c(prop_diff_est, w_est, rel_est, lv_est))




result <- ggplot(data = confints) + 
  geom_hline(aes(yintercept = .056), linetype = "dashed", color = "blue", alpha = .5) +
  geom_pointrange(aes(x = BootType, y = Est, ymin = `2.5%`, ymax = `97.5%`)) +
  annotate("segment", x = "Relative", y = .06, xend = "Relative", yend = .0565, 
           arrow = arrow(), size=.4, alpha=.5, color = "blue") + 
  annotate("text", x = "Relative", y = 0.064, 
           label = "True Difference", color = "blue", alpha = .5) +
  scale_y_continuous(breaks = c(.05, .056, .06, .07, .08, .09)) + 
  coord_flip() + 
  labs(y = "Estimated Prop. Diff. (Dem - Rep)", title = "Comparing estimations of 2018 U.S. Midterms")

result
```
